---
title: "Clustering"
output: html_notebook
---

```{r, echo=FALSE}
library(tidyverse)
library(readxl)
library(janitor) # for clean names
library(tidymodels)
library(ggpubr) # for adding regression line equations to scatter plots
library(skimr)
```

```{r project_functions}
source("../R_code/green_space_functions.R")
```

```{r}
theme_set(theme_light())
```

```{r notebook_functions}
normalise <- function(x){
  (x - min(x, na.rm = TRUE)) / (max(x, na.rm = TRUE) - min(x, na.rm = TRUE)) 
}
```

## Introduction

### Objectives

### The anticipated audience of the notebook

### Writing style

### The status of the notebook

### The datasets used

All data files are in the 'Data' folder (i.e. /foe_green_space/Data).

| File Name | Variable name in this notebook | Notes | More details |
|-----------|--------------------------------|-------|--------------|
|           |                                |       |              |
|           |                                |       |              |
|           |                                |       |              |

### A summary of findings / conclusions

### Other notes

## Ideas

-   Utility function for garden area per capita

-   Incorporating flats / no gardens

-   Rescaling variables

    -   range (0-1)
    -   variance?
    -   weighting?

**Preparing data for clustering:**

-   Read in data from excel
-   clean data
-   explore data (if needed)
-   transform data for clustering

[Preprocessing](https://medium.com/@evgen.ryzhkov/5-stages-of-data-preprocessing-for-k-means-clustering-b755426f9932):

-   Numerical variables only

-   Manage noise and outliers

-   Reduce skew (i.e. aim for symmetric distributions)

-   Place variables on the same scale

-   Check for collinearity

-   Minimize number of variables

[Check assumptions](https://www.r-bloggers.com/2017/08/exploring-assumptions-of-k-means-clustering-using-r/):

-   spherical clusters

-   approximately equal size clusters

-   approximately equal variance

## Prepare garden data for clustering

The ONS (Office for National Statistics) provides [data on access to private green space](https://www.ons.gov.uk/economy/environmentalaccounts/datasets/accesstogardensandpublicgreenspaceingreatbritain) (i.e. access to gardens) for each Local Authority District in Great Britain. Here I am using the most recent April 2020 edition of the data. I quickly, manually edited the ONS excel file to make it easier use the `read_excel` function for data import. Given it is unlikely that the ONS data will be updated during the course of this analysis, it was preferable to go for the quicker manual process than investing time in a re-producible programmatic approach.

**Notes on trying to understand how ONS figures are calculated:**

-   I was trying to work out how the ONS calculate `ave_gar_size` and `perc_add_with_gar`.

-   It is unclear how the average garden size (`ave_gar_size`) is calculated. It (`ave_gar_size_calc`) is not calculated as the total garden area divided by the number of addresses with a garden (`total_gar_area` / `add_with_gar_count`). On average the ratio of `ave_gar_size:ave_gar_size_calc`is greater than 1, indicating the ONS figure tends to be greater than my calculated figure. For now I will assume this is to do with differences in how flats and houses are treated in the calculation.

-   The percentage of addresses with a garden (`perc_add_with_gar)`is simply calculated as the number of addresses with a garden divided by the total numbers of addresses (i.e. `add_with_gar_count` / `add_count)`

**Notes on outliers:**

-   `ave_gar_size`: one clear outlier removed at around 120,000 m^2^

-   `perc_add_with_gar`: no obvious outliers as all values are between 0 and 1.

```{r import_gardens}

gardens <- read_ons_garden_data()

# focus on variables of interest
gardens_focus <- gardens %>% 
  
  # remove columns relating to house and flats, 
  # leaving columns relating to all households 
  # (i.e the total for both houses and flats)
  
  select(!(country_code:lad_name)) %>% 
  select(!(address_count_9:average_number_of_flats_sharing_a_garden)) %>% 
  rename(add_count = address_count_22,
         add_with_gar_count = adress_with_private_outdoor_space_count_23,
         total_gar_area = private_outdoor_space_total_area_m2_24,
         perc_add_with_gar = percentage_of_adresses_with_private_outdoor_space_25, 
         ave_gar_size = average_size_of_private_outdoor_space_m2_26)

# quickly exploring the data trying to understand how ONS figures are calculated
gardens_focus %>% 
  mutate(perc_add_with_gar_calc = add_with_gar_count / add_count,
         ons_calc_comp_perc = perc_add_with_gar/ perc_add_with_gar_calc,
         ave_gar_size_calc = total_gar_area / add_with_gar_count,
         ons_calc_comp_size = ave_gar_size / ave_gar_size_calc) %>% 
  summarise(mean_comp_perc = mean(ons_calc_comp_perc, na.rm = TRUE),
            mean_comp_size = mean(ons_calc_comp_size, na.rm = TRUE))

# focus down further on variables of interest
gardens_focus <- gardens_focus %>% 
  select(msoa_code, msoa_name, perc_add_with_gar, ave_gar_size)

# review variable distriubtions
p <- ggplot(data = gardens_focus)
p + geom_histogram(mapping = aes(x = perc_add_with_gar))
p + geom_histogram(mapping = aes(x = ave_gar_size))
p + geom_boxplot(mapping = aes(x = ave_gar_size)) +
  annotate(geom = "text", x = 10500, y = 0.1, 
           label = "Outlier with potential\nto impact on clustering\nperformance")

# remove outliers
gardens_focus <- gardens_focus %>% 
  filter(ave_gar_size < 3000)


# transform variables of interest to more normal distributions
gardens_trans <- gardens_focus %>% 
  mutate(log_ave_gar_size = log(ave_gar_size),
         log_reflect_perc_add_with_gar = 1 / (max(perc_add_with_gar + 1, na.rm = TRUE) - perc_add_with_gar + 1),
         exp_perc_add_with_gar = 10**(perc_add_with_gar),
         to_power_six_perc_add_with_gar = perc_add_with_gar**6) 
  

# Create dataframe of scaled and centred garden data
gardens_scaled <- gardens_trans %>%
  mutate(across(.cols = where(is.numeric), 
                .fns = scale,
                .names = "scale_{.col}"
                )
         )

# Create dataframe of normalised data
gardens_norm <- gardens_trans %>%
  mutate(across(.cols = where(is.numeric), 
                .fns = normalise,
                .names = "norm_{.col}"
                )
         )

# plot scaled data
p <- ggplot(data = gardens_scaled)

p + geom_histogram(mapping = aes(x = scale_to_power_six_perc_add_with_gar)) 
p + geom_histogram(mapping = aes(x = scale_log_ave_gar_size))

# plot normalised data
p <- ggplot(data = gardens_norm)

p + geom_histogram(mapping = aes(x = norm_to_power_six_perc_add_with_gar)) 
p + geom_histogram(mapping = aes(x = norm_log_ave_gar_size))

```

## Prepare public green space data for clustering

**Notes on outliers:**

-   `green_space_area_per_capita`: no obvious outliers on boxplot with log10 scale.

-   `pcnt_pop_with_go_space_access`: no obvious outliers as all values are between 0 and 1.

```{r}
# read in data from excel
foe_green_space <- read_foe_data()

# focus on variables of interest
foe_green_space_focus <- foe_green_space %>% 
  select(msoa_code, msoa_name, green_space_area_per_capita, pcnt_pop_with_go_space_access)

# review variable distriubtions
p <-  ggplot(data = foe_green_space_focus)
p + geom_histogram(mapping = aes(x = green_space_area_per_capita))
p + geom_boxplot(mapping = aes(x = green_space_area_per_capita)) + scale_x_log10()
p + geom_histogram(mapping = aes(x = pcnt_pop_with_go_space_access))

# no obvious outliers to remove
  
# transform variables of interest to more normal distributions
foe_green_space_trans <- foe_green_space_focus %>% 
  mutate(log_green_space_area_per_capita = log(green_space_area_per_capita + 1),
         sqrt_pcnt_pop_with_go_space_access = sqrt(pcnt_pop_with_go_space_access)
         )

# Create dataframe of scaled and centred garden data
foe_green_space_scaled <- foe_green_space_trans %>% 
  mutate(across(.cols = where(is.numeric), 
                .fns = scale,
                .names = "scale_{.col}"
                )
         )

# Create dataframe of normalised data
foe_green_space_norm <- foe_green_space_trans %>% 
  mutate(across(.cols = where(is.numeric), 
                .fns = normalise,
                .names = "norm_{.col}")
         )
  
# plot scaled data
p <- ggplot(data = foe_green_space_scaled)
p + geom_histogram(mapping = aes(x = scale_log_green_space_area_per_capita))
p + geom_histogram(mapping = aes(x = scale_sqrt_pcnt_pop_with_go_space_access))

# plot normalised data
p <- ggplot(data = foe_green_space_norm)
p + geom_histogram(mapping = aes(x = norm_log_green_space_area_per_capita))
p + geom_histogram(mapping = aes(x = norm_sqrt_pcnt_pop_with_go_space_access))
```

## Merge gardens and green space datasets

```{r}
gs_gar_scaled <- foe_green_space_scaled %>% 
  left_join(gardens_scaled)

gs_gar_norm <- foe_green_space_norm %>% 
  left_join(gardens_norm)  
  
```

## Pre-process the green space - gardens datasets

```{r}
# select (normalised) variables for use in clustering
data_points_norm <- gs_gar_norm %>% 
  select(msoa_code, msoa_name,
         norm_log_green_space_area_per_capita, 
         norm_sqrt_pcnt_pop_with_go_space_access,
         norm_log_ave_gar_size,
         norm_to_power_six_perc_add_with_gar
         ) %>%   
  
  # remove any nas
  filter(across(.cols = where(is.numeric),
                .fns = ~ !is.na(.x)))

# select (scaled) variables for use in clustering
data_points_scaled <- gs_gar_scaled %>% 
  select(scale_log_green_space_area_per_capita, 
         scale_sqrt_pcnt_pop_with_go_space_access,
         scale_log_ave_gar_size,
         scale_to_power_six_perc_add_with_gar
         ) %>%   
  
  # remove any nas
  filter(across(.cols = where(is.numeric),
                .fns = ~ !is.na(.x)))
```

**Check for collinearity between variables**

-   Only conducted for normalized data, would expect the same correlations for scaled data.

-   Highest correlation coefficient (0.55): `norm_log_ave_gar_size` - `norm_log_green_space_area_per_capita`

-   Second highest correlation coefficient (0.42) `norm_log_ave_gar_size` - `norm_sqrt_pcnt_pop_with_go_space_access`

-   Neither of the above would be considered strong correlations (by the benchmark of Pearson's correlation coefficient being 0.7 or greater.

-   Arguably the relationship between `norm_log_ave_gar_size` - `norm_log_green_space_area_per_capita` could be considered a moderate correlation, as the correlation coefficient is greater than 0.5. I think this degree of correlation could indicate that both variables are related to (population) density. In other words, denser MSOAs have both smaller gardens and less public green space per capita. For now, I'll retain both variables and reconsider if either could or should be omitted from clustering based on the initial results.

**Minimize number of variables**

Given the lack of strong correlations between the four variables, the four variables selected is the minimum required to incorporate information on both:

-   The amount and accessibility of public green space;

-   And, the amount and accessibility of private green space;

```{r fig.height=5}
library(corrr)

# identify correlations between the variables
d <- correlate(data_points_norm %>% select(where(is.numeric)))
d
d %>% 
  rplot() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

# plot pairs of variables where correlations have been indentified
p <- ggplot(data = data_points_norm)

p + geom_point(mapping = aes(x = norm_log_green_space_area_per_capita, norm_log_ave_gar_size))

p + geom_point(mapping = aes(x = norm_sqrt_pcnt_pop_with_go_space_access, norm_log_ave_gar_size))
```

Looking at the relationship between `norm_log_ave_gar_size`, `norm_log_green_space_area_per_capita` and `population_density` in more detail. From r-squared values from the linear regression models plotted below, we can observe that:

-   Approximately two thirds of the variability in `norm_log_ave_gar_size` can be explained by the natural logarithm of `population_density`;
-   Approximately half in `norm_log_green_space_area_per_capita` can be explained by the natural logarithm of `population_density`;

```{r}
# add population density data to the two variables of interest
pop_den_gs <- foe_green_space %>% 
  select(msoa_code, msoa_name, population_density) %>% 
  right_join(data_points_norm) %>% 
  select(population_density, norm_log_ave_gar_size, norm_log_green_space_area_per_capita) %>% 
  mutate(log_population_density = log(population_density)) %>% 
  pivot_longer(cols = starts_with("norm"), names_to = "variable", values_to = "value")


# Plot the two variables vs population density
p <-  ggplot(data = pop_den_gs,
             mapping = aes(x = log_population_density,
                           y = value,
                           colour = variable)
)

p +  geom_point() +
  
  stat_smooth(method = "lm", colour = "grey50") +
  
  stat_regline_equation(colour = "grey50",
                        label.x = -7.5) +
  
  stat_cor(mapping = aes(label = paste(..rr.label.., ..p.label.., sep = "~`,`~")), # add R2 value to plot
           colour = "grey50", 
           label.x = -8.0,
           label.y = 0.9) +
  
  facet_wrap(~variable) +
  
  theme(legend.position = "none")
```

## Clustering with normalised data

```{r}

```

Run kmeans algorithm once to understand the function and its output

```{r}
# final preparation of data for k means algorithm 
# remove non-numeric (key) columns
kmeans_in <- data_points_norm %>% 
  select(-starts_with("msoa"))

# define number of clusters
k <- 3

# run k-means algorithm
kmeans_out <- kmeans(kmeans_in,
                     centers = k)

# display k-means algorithm output
augment(kmeans_out, data_points_norm)
tidy(kmeans_out)
glance(kmeans_out)

```

Run exploratory clustering - looking at the choice of k

```{r}

run_n_kmeans <- function(n = 9, km_in, data){

  kmeans_outputs <- tibble(k = 1:n) %>% 
    mutate(kclust = map(k, ~ kmeans(km_in, centers = .x)),
           tidied = map(kclust, tidy),
           glanced = map(kclust, glance),
           augmented = map(kclust, augment, data = data)
           )
  
  return(kmeans_outputs)
}


get_n_kmeans_outputs <- function(kmeans_outputs_df, column_name){
  kmeans_outputs_df %>% 
    unnest(cols = c({{column_name}}))
}

plot_kmeans_clusters <- function(assignments, clusters, x, y) {
  ggplot(data = assignments,
             mapping = aes(x = !!sym(x),
                           y = !!sym(y))
             ) +
    
  geom_point(mapping = aes(colour = .cluster), alpha = 0.8) +
    
  geom_point(data = clusters,
             size = 5,
             shape = "x") +
  scale_colour_viridis_d() +
    
  facet_wrap(~ k)
}

plot_kmeans_summary_stat <- function(clusterings_df, summary_stat){
  ggplot(data = clusterings_df,
       mapping = aes(x = k, y = {{summary_stat}})) +
  geom_line() +
  geom_point()
}


```

```{r}
# create a class for objects which store exploratory workflow outputs
setClass(Class = "kmeans_wf_output",
         representation(clusters = "data.frame",
                        assignments = "data.frame",
                        clusterings = "data.frame")
         )

# define the exploratory workflow
run_kmeans_exploratory_workflow <- function(km_in, data, k_max, public_gs_vars, garden_vars){
  
  # run k means algorithm
  kmeans_outputs <- run_n_kmeans(n = k_max, km_in = km_in, data = data)
  
  # process kmeans outputs
  clusters <- get_n_kmeans_outputs(kmeans_outputs, tidied) %>% 
    select(-kclust, -augmented, -glanced) # remove columns of large-ish tibbles
  assignments <- get_n_kmeans_outputs(kmeans_outputs, augmented) %>% 
    select(-kclust, -tidied, -glanced)
  clusterings <- get_n_kmeans_outputs(kmeans_outputs, glanced) %>% 
    select(-kclust, -tidied, -augmented)
  
  # produce plots for visual inspection
  plot(plot_kmeans_clusters(assignments, clusters,
                           public_gs_vars[[1]],
                           public_gs_vars[[2]]
                           )
      )

  plot(plot_kmeans_clusters(assignments, clusters,
                            garden_vars[[1]],
                            garden_vars[[2]]
                            )
       )

  plot(plot_kmeans_summary_stat(clusterings, tot.withinss)
       )
  
  # return dataframes created by the workflow
  res <- new("kmeans_wf_output", 
             assignments = assignments, clusters = clusters, clusterings = clusterings)
  return(res)
}

# final preparation of data for k means algorithm 
# remove non-numeric (key) columns
kmeans_in_norm <- data_points_norm %>% 
  select(-starts_with("msoa"))

kmeans_in_scale <- data_points_scaled %>% 
  select(-starts_with("msoa"))

# define variables for scatter plots (used within the exploratory workflow)
public_gs_vars_norm <- c("norm_log_green_space_area_per_capita",
                    "norm_sqrt_pcnt_pop_with_go_space_access")

garden_vars_norm <- c("norm_log_ave_gar_size",
                 "norm_to_power_six_perc_add_with_gar")

public_gs_vars_scale <- c("scale_log_green_space_area_per_capita",
                    "scale_sqrt_pcnt_pop_with_go_space_access")

garden_vars_scale <- c("scale_log_ave_gar_size",
                 "scale_to_power_six_perc_add_with_gar")

# run workflows for normalised and scaled data
exploratory_workflow_norm_out <- run_kmeans_exploratory_workflow(kmeans_in_norm, data_points_norm, public_gs_vars_norm, garden_vars_norm, k_max = 9)

exploratory_workflow_scale_out <- run_kmeans_exploratory_workflow(kmeans_in_scale, data_points_scaled, public_gs_vars_scale, garden_vars_scale, k_max = 9)
```

**k = 5:**

```{r}
# get workflow outputs (normed data)
norm_assignments <- exploratory_workflow_norm_out@assignments

# add in unprocessed garden and public green space data
gs_gar_norm_clusters <- norm_assignments %>% 
  left_join(gs_gar_norm)


# get workflow outputs (scaled data)
scale_assignments <- exploratory_workflow_scale_out@assignments

# add in unprocessed garden and public green space data
gs_gar_scale_clusters <- scale_assignments %>% 
  left_join(gs_gar_scaled)
```

```{r}
library(ggridges)

cluster_ridge_plots <- function(df, var_of_int, k_of_int = c(4,5)){

  p <- ggplot(data = df %>% filter(k %in% k_of_int),
              mapping = aes(x = {{var_of_int}},
                            y = .cluster,
                            fill = .cluster)
              ) +
    geom_density_ridges() +
    scale_fill_viridis_d() +
    facet_wrap(~k) +
    theme_light()
  
  return(p)
}

cluster_ridge_plots(gs_gar_norm_clusters, perc_add_with_gar)
cluster_ridge_plots(gs_gar_norm_clusters, ave_gar_size)
cluster_ridge_plots(gs_gar_norm_clusters, pcnt_pop_with_go_space_access)
cluster_ridge_plots(gs_gar_norm_clusters, log_green_space_area_per_capita)


```

```{r}

plot_cluster_ridges <- function(k_int, clusters_df){
  plotting_df_clusters <- clusters_df %>% 
    filter(k == k_int) %>% 
    select(msoa_code, msoa_name,
           perc_add_with_gar, log_ave_gar_size,
           pcnt_pop_with_go_space_access, log_green_space_area_per_capita,
           .cluster) %>% 
    pivot_longer(cols = perc_add_with_gar:log_green_space_area_per_capita,
                 names_to = "green_space_variable",
                 values_to = "values")
  
  ggplot(data = plotting_df_clusters,
         mapping = aes(x = values,
                       y = .cluster,
                       fill = .cluster)
  ) +
    geom_density_ridges() +
    scale_fill_viridis_d() +
    facet_wrap(~green_space_variable, scales = "free_x")

}

plot_cluster_ridges(k_int = 6, clusters_df = gs_gar_norm_clusters)
plot_cluster_ridges(k_int = 5, clusters_df = gs_gar_scale_clusters)



# gs_gar_scale_clusters %>% 
#   select(msoa_code, msoa_name,
#          scale_to_power_six_perc_add_with_gar, scale_log_ave_gar_size,
#          scale_sqrt_pcnt_pop_with_go_space_access, scale_log_green_space_area_per_capita) %>% 
#   skim()

```

## Interpreting clusters (scaled)

I think this makes more sense as normalisation brings points closer to the mean, but there data points distant from the mean (e.g. MSOA's with low percentages of people with gardens / green space access) that should be receive weighting in the context of the focus on green space deprivation.

From the plot of total sum of squares within clusters above, I think it is worth considering k = 4, 5 and 6 in more detail.

### k = 4

```{r}
calc_cluster_means <- function(df, k_of_int){
  df %>%
    select(k, msoa_code, msoa_name,
           ave_gar_size, green_space_area_per_capita,
           pcnt_pop_with_go_space_access, perc_add_with_gar,
           .cluster) %>% 
    filter(k == k_of_int) %>% 
    group_by(.cluster) %>% 
    summarise(n = n(),
              across(where(is.numeric), ~mean(.x))) %>% 
    arrange(desc(.cluster))
}

k_of_int = 4

plot_cluster_ridges(k_int = k_of_int, clusters_df = gs_gar_scale_clusters)
calc_cluster_means(gs_gar_scale_clusters, k_of_int)
```

+---+---------+-------------------+----------------+-------------------------------+-----------------------------+
| k | Cluster | perc_add_with_gar | ave_gar_size   | pcnt_pop_with_go_space_access | green_space_area_per_capita |
+===+=========+===================+================+===============================+=============================+
| 4 | 1       | High (rel)        | Medium (rel)   | High (rel)                    | medium(rel)                 |
|   |         |                   |                |                               |                             |
|   |         | High (abs)        | Good (abs)     | Med-high (abs)                | adequate(abs)               |
+---+---------+-------------------+----------------+-------------------------------+-----------------------------+
| 4 | 2       | Low (rel)         | low-med(rel)   | low-high(rel)                 | low-med(rel)                |
|   |         |                   |                |                               |                             |
|   |         | Med-high(abs)     | poor-adeq(abs) | poor-good(abs)                | poor-ade(abs)               |
+---+---------+-------------------+----------------+-------------------------------+-----------------------------+
| 4 | 3       | High (rel)        | High(rel)      | low-med(rel)                  | med-good(rel)               |
|   |         |                   |                |                               |                             |
|   |         | High (abs)        | Good(abs)      | poor-adeq(abs)                | good-exce(abs)              |
+---+---------+-------------------+----------------+-------------------------------+-----------------------------+
| 4 | 4       | High (rel)        | Medium (rel)   | low-med(rel)                  | low-med(rel)                |
|   |         |                   |                |                               |                             |
|   |         | High (abs)        | Good (abs)     | poor-adeq(abs)                | poor-adeq(abs)              |
+---+---------+-------------------+----------------+-------------------------------+-----------------------------+

### k = 5

```{r}
k_of_int = 5

plot_cluster_ridges(k_int = k_of_int, clusters_df = gs_gar_scale_clusters)
calc_cluster_means(gs_gar_scale_clusters, k_of_int)
```

+---+---------+---------------------+-----------------------------+-------------------------------+-------------------+
| k | Cluster | ave_gar_size        | green_space_area_per_capita | pcnt_pop_with_go_space_access | perc_add_with_gar |
+===+=========+=====================+=============================+===============================+===================+
| 5 | 5       | **medium**          | **medium**                  | **med-high**                  | **med-high**      |
|   |         |                     |                             |                               |                   |
|   |         | adeq-good           | adeq                        | *poor-adeq*                   | *poor-good*       |
+---+---------+---------------------+-----------------------------+-------------------------------+-------------------+
| 5 | 4       | **medium**          | **medium**                  | **med-high**                  | **high**          |
|   |         |                     |                             |                               |                   |
|   |         | adeq-good           | adeq                        | *poor-adeq*                   | *good*            |
+---+---------+---------------------+-----------------------------+-------------------------------+-------------------+
| 5 | 3       | **\*\#low**         | **\*\#low**                 | **low-med**                   | **med-high**      |
|   |         |                     |                             |                               |                   |
|   |         | *poor-adeq***\#\*** | *poor-adeq***\#\***         | *poor*                        | *poor-good*       |
+---+---------+---------------------+-----------------------------+-------------------------------+-------------------+
| 5 | 2       | **v high**          | **v high**                  | **low-med**                   | **high**          |
|   |         |                     |                             |                               |                   |
|   |         | *excellent*         | *excellent*                 | *poor*                        | *good*            |
+---+---------+---------------------+-----------------------------+-------------------------------+-------------------+
| 5 | 1       | **medium**          | **\*\#low**                 | **low-med**                   | **high**          |
|   |         |                     |                             |                               |                   |
|   |         | adeq-good           | *poor-adeq***\#\***         | *poor*                        | *good*            |
+---+---------+---------------------+-----------------------------+-------------------------------+-------------------+

### k = 6

```{r}
k_of_int = 6

plot_cluster_ridges(k_int = k_of_int, clusters_df = gs_gar_scale_clusters)
calc_cluster_means(gs_gar_scale_clusters, k_of_int)
```

```{r}
p2_public_gs <- p1_public_gs +
  geom_point(data = clusters,
             size = 5,
             shape = "x")
p2_public_gs

p2_gardens <- p1_gardens +
  geom_point(data = clusters,
             size = 5,
             shape = "x")
p2_gardens
```

```{r}

```

```{r eval=FALSE}

# could create a reflect function
%>% 
  mutate(reflect_perc_add_with_gar = max(perc_add_with_gar, na.rm = TRUE) - perc_add_with_gar)

p <-  ggplot(data = gardens,
             mapping = aes(x = reflect_perc_add_with_gar))

p + geom_histogram() + scale_x_log10()
```

```{r}
# clusters <- kmeans_outputs %>% 
#   unnest(cols = c(tidied)) %>% 
#   select(-kclust, -augmented, -glanced)
# clusters
# 
# assignments <- kmeans_outputs %>% 
#   unnest(cols = c(augmented)) %>% 
#   select(-kclust, -tidied, -glanced)
# assignments
# 
# clusterings <- kmeans_outputs %>% 
#   unnest(cols = c(glanced)) %>% 
#   select(-kclust, -tidied, -augmented)
# 
# clusterings
```
